---
title: "Untitled"
author: "Madeline Egbert"
date: "03/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyverse)
```

```{r}
library(readxl)
defense <- read_excel('cyclonesFootball2020.xlsx', sheet = 'Defensive')
str(defense)

```
```{r}
offense <- read_excel('cyclonesFootball2020.xlsx', sheet = 'Offensive')
str(offense)

```
```{r}
biography <- read_excel('cyclonesFootball2020.xlsx', sheet = 'Biography')
str(biography)

```

#Part1
#Cleaning Data - Manipulate the three imported data frames, so that in the cleaned dataset:

#1: The player names and opponent names are factors.
```{r}
Bio <- biography %>%
  mutate(Name = as.factor(Name))
str(Bio)
```

```{r}
offClean <- offense %>%
  mutate(Name = as.factor(Name)) %>%
  mutate(Opponent = as.factor(Opponent_Opponent))
offClean = subset(offClean, select = -c(Opponent_Opponent) )
str(offClean)
```
```{r}
defClean <- defense %>%
  mutate(Name = as.factor(Name)) %>%
  mutate(Opponent = as.factor(Opponent_Opponent))
defClean = subset(defClean, select = -c(Opponent_Opponent) )
str(defClean)
```

#2: All offensive and defensive statistics are numerical. The Weight column in biogrpahy is also numeric.
```{r}
offClean <- offClean %>% 
  mutate_if(is.character, as.numeric)
```
```{r}
defClean <- defClean %>% 
  mutate_if(is.character, as.numeric)
```
```{r}
Bio <- Bio %>%
  mutate(Weight = as.numeric(Weight))
```

#3: Change the Height column in biography into numeric. (Hint: Use a unit that works well
```{r}
bioClean <- Bio %>%
  separate(Height,c('feet', 'inches'), sep = '-', convert = TRUE, remove = FALSE) %>%
  mutate(feet = 12*feet + inches) %>%
  select(-inches) %>%
  rename(Height_In_Inches = feet) %>%
  mutate(Height_In_Inches = as.numeric(Height_In_Inches))
bioClean = subset(bioClean, select = -c(Height) )
str(bioClean)
```


#Part2

#1
```{r}
offClean1 <- offClean %>%
  pivot_longer(Receiving_REC:Passing_INT, names_to = 'stat', values_to = 'value')
str(offClean1)
```
#2
```{r}
ggplot(offClean1, aes(x=value)) + geom_histogram(binwidth= 30) + facet_wrap(~stat, scales= 'free_y')
```
### A lot of the histograms show that they have null values for the offensive stats. Only the Passing_YDS, Receiving_YDS, Rushing_ATT, and Rushing_YDS have varied distributions. All of these are skewed to the left, and Passing_YDS has some outliers. 

#3





#4
```{r}
CityState <- bioClean %>%
  separate(Hometown, c('City', 'State'), sep = ',', convert = TRUE, remove = FALSE) %>%
  select(-Position, -Height_In_Inches, -Weight, -Class, -Hometown, -Highschool)
head(CityState)
```
#5
```{r}
PlayersPerState <- CityState %>%
  group_by(State) %>%
  summarize(count = n())
PlayersPerState
```


#Part3


#1
```{r}
offState <- offClean %>%
  left_join(CityState, by = c('Name')) %>%
  select(Name, State) %>%
  distinct()
offState
```
```{r}
ggplot(offState, aes(x = Name, y = State)) + geom_point() + theme(axis.text.x = element_text(angle = 90)) + ggtitle('ISU Offense vs. State Origin')
```



```{r}
defState <- defClean %>%
  left_join(CityState, by = c('Name')) %>%
  select(Name, State) %>%
  distinct()
defState
```

```{r}
ggplot(defState, aes(x = Name, y = State)) + geom_point() + theme(axis.text.x = element_text(angle = 90)) + ggtitle('ISU Defense vs. State Origin')
```

#2


#3

```{r}
def2019 <- read_excel('cyclonesFootball2019.xlsx', sheet = 'Defensive')
str(def2019)

```
```{r}
off2019 <- read_excel('cyclonesFootball2019.xlsx', sheet = 'Offensive')
str(off2019)

```

```{r}
bio2019 <- read_excel('cyclonesFootball2019.xlsx', sheet = 'Biography')
str(bio2019)
```


```{r}
off2019Clean <- off2019 %>%
  mutate(Name = as.factor(Name)) %>%
  mutate(Opponent = as.factor(Opponent_Opponent))
off2019Clean = subset(off2019Clean, select = -c(Opponent_Opponent))
str(off2019Clean)
```

```{r}
def2019Clean <- def2019 %>%
  mutate(Name = as.factor(Name)) %>%
  mutate(Opponent = as.factor(Opponent_Opponent))
def2019Clean = subset(def2019Clean, select = -c(Opponent_Opponent))
str(def2019Clean)
```

```{r}
off2019Clean <- off2019Clean %>%
  mutate_if(is.character, as.numeric)
```

```{r}
def2019Clean <- def2019Clean %>%
  mutate_if(is.character, as.numeric)
```

```{r}
str(off2019Clean)
```

```{r}
str(def2019Clean)
```

```{r}
def2019Clean <- def2019Clean %>%
  group_by(Name) %>%
  summarize_at(vars(Tackles_Solo), list(Total_Solo_Tackles2019 = sum, Mean_Solo_Tackles2019 = mean))
def2019Clean
```





```{r}
off2019Clean <- off2019Clean %>%
  group_by(Name) %>%
  summarize_at(vars(Receiving_YDS), list(Total_Receiving_YDS_2019 = sum, Mean_Receiving_YDS2019 = mean))
off2019Clean
```




```{r}
defprogress2019_to_2020 <- def2019Clean %>% inner_join(defClean)
defprogress2019_to_2020
```


```{r}
offprogress2019_to_2020 <- off2019Clean %>% 
  inner_join(offClean)
offprogress2019_to_2020
```


